j_play

### 2020-12-07 scripts/Final_project.Rmd
References push for inline code

bibliography      : "r-references.bib"

@@

# Introduction
Educational policy can impact practice by reorganizing educational delivery and reconceptualizing the meaning and impact of education on society (Ball, 1998). Educational policy exists as a mechanism to drive changes in school buildings that result in socially important outcomes (e.g., increasing literacy acquisition through improved instruction, reducing instances of bullying through prevention programs). However, the relationship between policy and practice in Western formal systems of education is often unclear, misdirected, or atheoretical (Cohen et al., 2009). Leveraging the power of policy as a change agent for the delivery and conceptualization of education stands to benefit a significant proportion of students (Swanson, 2007).

When policy is enacted, the intended outcomes are often not realized for months, years, or decades (Gu et al., 2018). For example, changing school district policy on professional development to align with research-informed practices may be constructed to increase student academic achievement; though, there is insufficient evidence pointing to how long professionalized intervention with teachers will take to increase student academic scores (Luke & McArdle, 2009). Longitudinal data helps to understand the impact of policy change because of the year-over-year comparisons that illustrate differences, inflection points, or anomalies of interest (Michelmore & Dynarski, 2017).

@@

We intend to explore several questions regarding longitudinal trends where educational policy may have impacted special education practice. First, we will use descriptive statistics to describe trends in disability prevalence over time. The initial analyses will inform a discussion on developmental trajectories by illustrating shifts in prevalence growth from static to linearly increasing based on age of diagnosis for the Oregon data, which can serve as a springboard to make inferences about BC data (e.g., common age of onset for diagnosis per category and diagnostic patterns over time). In studying the differences and similarities between OR and BC datasets, we will also engage in a discussion on diagnostic terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders (DSM; American Psychological Association, 2013), in part as a response to a challenge set by differing terminology between BC and OR categorization (i.e., the United States federal guidelines for exceptionality designation are similar but not identical to exceptionality designation guidelines in BC).

Second, we hope to closely analyze any changes, or lack thereof, within the BC data following Canadian Supreme Court rulings between the British Columbia Teachers Federation and the province of BC (See Lambert, 2017). Canadian Supreme Court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers’ that school districts are required to hire. The driving theory of change is that decreasing student-to-teacher ratios and capping the number of students with exceptionalities in each classroom causes student designation practices to change. Theoretically, school district administrators made decisions based on current staffing levels, student designations, diagnostic capacity and malleability, as well as external pressures to adjust practice according to legal mandates.

@@

## Problem Statement and RQs
In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker & Bergmann, 2013). While Canadian K-12 academic achievement outcomes are generally higher than a national-level comparison with the US, there are ongoing policy issues to address when provincial ministries of education are crafting legislation and procedures to impact a top-tier system of education. Significant events, such as the 2016 Canadian supreme court ruling that directly impacted British Columbia classrooms, have downstream effects on instructional interactions; yet a retroactive policy lens is rarely applied after such events occur. The current study seeks to address the knowledge gap of downstream instructional interact effects after a significant event in BC educational policy.

@@

The rapid decrease in students identified as gifted in rural districts between the years of 2004 and 2007 cannot be explained by one sole factor in this exploratory analysis. With respect to the speculative nature of this assertion, the authors assume that educational policy at the provincial level impacts educational practice across school districts. In 2006, the British Columbia Ministry of Education released Special education services: A manual of policies, procedures and guidelines, which explicitly states the acceleration programming that should be afforded to students with exceptional academic talent (Kanevsky & Clelland, 2016). In fact, this 2006 version of the manual was a revision of the original 1995 manual, and this initial version included scant guidance on service provision to gifted students. This 2016 shift in policy links to a higher degree of resources afforded to gifted students through “independent, guided education” (p. 52). Given the additional resources needed to properly implement gifted educational services, rural school districts may have de-designated students and sought other ways to provide instruction that meets the needs of their students. For example, Lo et al. (2019) suggest that the British Columbia Ministry of Education’s recent curriculum update makes amends for the traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.

>>>>

bibliography      : ["r-references.bib"]

@@

# Introduction
Educational policy can impact practice by reorganizing educational delivery and reconceptualizing the meaning and impact of education on society [@ball_1998]. Educational policy exists as a mechanism to drive changes in school buildings that result in socially important outcomes (e.g., increasing literacy acquisition through improved instruction, reducing instances of bullying through prevention programs). However, the relationship between policy and practice in Western formal systems of education is often unclear, misdirected, or atheoretical [@cohen_2009]. Leveraging the power of policy as a change agent for the delivery and conceptualization of education stands to benefit a significant proportion of students [@swanson_2007].

When policy is enacted, the intended outcomes are often not realized for months, years, or decades [@gu-2018]. For example, changing school district policy on professional development to align with research-informed practices may be constructed to increase student academic achievement; though, there is insufficient evidence pointing to how long professionalized intervention with teachers will take to increase student academic scores [@luke_2009]. Longitudinal data helps to understand the impact of policy change because of the year-over-year comparisons that illustrate differences, inflection points, or anomalies of interest [@michelmore_2017].

@@

We intend to explore several questions regarding longitudinal trends where educational policy may have impacted special education practice. First, we will use descriptive statistics to describe trends in disability prevalence over time. The initial analyses will inform a discussion on developmental trajectories by illustrating shifts in prevalence growth from static to linearly increasing based on age of diagnosis for the Oregon data, which can serve as a springboard to make inferences about BC data (e.g., common age of onset for diagnosis per category and diagnostic patterns over time). In studying the differences and similarities between OR and BC datasets, we will also engage in a discussion on diagnostic terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders [DSM, @american_2013], in part as a response to a challenge set by differing terminology between BC and OR categorization (i.e., the United States federal guidelines for exceptionality designation are similar but not identical to exceptionality designation guidelines in BC).

Second, we hope to closely analyze any changes, or lack thereof, within the BC data following Canadian Supreme Court rulings between the British Columbia Teachers Federation and the province of BC [see @lambert_2017]. Canadian Supreme Court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers’ that school districts are required to hire. The driving theory of change is that decreasing student-to-teacher ratios and capping the number of students with exceptionalities in each classroom causes student designation practices to change. Theoretically, school district administrators made decisions based on current staffing levels, student designations, diagnostic capacity and malleability, as well as external pressures to adjust practice according to legal mandates.

@@

## Problem Statement and RQs
In comparison to the United States, Canadian education policy receives little attention and scholarly interest [@walker_2013]. While Canadian K-12 academic achievement outcomes are generally higher than a national-level comparison with the US, there are ongoing policy issues to address when provincial ministries of education are crafting legislation and procedures to impact a top-tier system of education. Significant events, such as the 2016 Canadian supreme court ruling that directly impacted British Columbia classrooms, have downstream effects on instructional interactions; yet a retroactive policy lens is rarely applied after such events occur. The current study seeks to address the knowledge gap of downstream instructional interact effects after a significant event in BC educational policy.

@@

The rapid decrease in students identified as gifted in rural districts between the years of 2004 and 2007 cannot be explained by one sole factor in this exploratory analysis. With respect to the speculative nature of this assertion, the authors assume that educational policy at the provincial level impacts educational practice across school districts. In 2006, the British Columbia Ministry of Education released Special education services: A manual of policies, procedures and guidelines, which explicitly states the acceleration programming that should be afforded to students with exceptional academic talent [@kanevsky_2013]. In fact, this 2006 version of the manual was a revision of the original 1995 manual, and this initial version included scant guidance on service provision to gifted students. This 2016 shift in policy links to a higher degree of resources afforded to gifted students through “independent, guided education” (p. 52). Given the additional resources needed to properly implement gifted educational services, rural school districts may have de-designated students and sought other ways to provide instruction that meets the needs of their students. For example, @lo_2019 suggest that the British Columbia Ministry of Education’s recent curriculum update makes amends for the traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
